Imagine you are a member of a jury judging a driving accident
case.  A taxi was seen to back into a parked car one night, creasing
the car's bodywork. The
entire case against the taxi company rests on the evidence of one
witness, an man who saw the accident from his apartment window.
He says that he saw a blue 
taxi backing into the car.  In trying to establish her case, the
lawyer for the 
car owner establishes the following facts:
\begin{itemize}
\item There are only two taxi companies in town, ``Blue Cabs'' and
  ``Green Cabs.''  On the night in question, 85\% of all taxis on the
  road were green and 15\% were blue.
\item The witness has undergone a third-party vision test under
  conditions similar to those on the night in question, and has
  demonstrated that he can successfully distinguish a blue taxi from a
  green taxi 80\% of the time.
\end{itemize}

The standard for proof in civil cases (as opposed to criminal cases)
is ``preponderance of evidence,'' often taken to mean that the claim
against the defendant is more likely than not.

Based on the information given above, how confident are you, the
juror, in ruling in favor of the car owner collecting reasonable
damages from the blue taxi company?

\begin{MultipleChoice}[itemname=before]
\wrong{Strongly against the ruling}
\correct{Against the ruling}
\wrong{Neutral}
\wrong{For the ruling}
\wrong{Strongly for the ruling}
\end{MultipleChoice}


Now do a calculation.   Construct a tree diagram with 100,000
hypothetical cases. 
\begin{enumerate}
\item  Divide this into 85,000 green and 15,000 blue
taxis.  
\item Further divide each branch into what the witness saw:
blue or green.  Use the 80 percent accuracy rate for the witness in
filling this in. 
\item You should have four end-points.  Use these to calculate the
  probability that the taxi was blue {\em given} that the witness saw
  blue.
\end{enumerate}

\begin{itemize}
\item What is the probability of the taxi being blue, given that the
witness saw blue?
\SelectSetHoriz[itemname=prob]{0.41}{0.15,0.17,0.29,0.41,0.68,0.71,0.80,0.85}
\end{itemize}

Based on this calculated probability, how how confident are you, the
juror, in ruling in favor of the car owner collecting reasonable
damages from the blue taxi company?  (Please do not change 
your previous answer to this question.) 

\begin{MultipleChoice}[itemname=after]
\wrong{Strongly against the ruling}
\correct{Against the ruling}
\wrong{Neutral}
\wrong{For the ruling}
\wrong{Strongly for the ruling}
\end{MultipleChoice}

\bigskip

\noindent{\small Source: From Utts (2010), ``Unintentional lies in the
  media: don't blame journalists for what we don't teach'' ICOTS8, who
  drew from  Tversky and
  Kahneman (1982), ``Evidential impact of base rates'' in {\em
    Judgment under uncertainty: Heuristics and biases''}, pp. 153-160 }

