\begin{enumerate}[(a)]
\item Why did you take Epi? (Mark all that apply.)
\begin{MultipleChoice}
\wrong{Planning a concentration in CGH}
\wrong{Considering medical school}
\wrong{Considering a career in public health}
\wrong{Just curious}
\wrong{Fulfill the QT requirement}
\wrong{Fulfill the science requirement}
\end{MultipleChoice}]

Other? \TextEntry

\vspace*{.5in}

\item Please comment on the choice of topics in the course, particularly on those you found particularly helpful and illuminating or those you found not helpful.  To remind you, the basic outline was
\begin{itemize}
\item Measuring health and disease (e.g. prevalence, risk ratio)
\item Outbreak and spread of disease
\item Cause and effect
\item Risk
\item Detection
\item Trade-Offs and health-care reform.
\end{itemize}

\vspace*{2in}

\item Is there an additional topic or topics you would like to see covered in this course?

\vspace*{1in}

\item How helpful was the textbook?

\vspace*{1in}

\item How helpful were the non-textbook book readings?

\vspace{1in}

\item How helpful were the article readings?

\vspace*{1in}

\item How helpful were the Epiville exercises and Island project?

\vspace*{1in}

\item Did you get what you expected out of this course?

\vspace*{1in}

\item How could the course be improved?

\vspace*{1in}

\item The class is quite large compared to most at Macalester.  How do you think this affected your experience?

\vspace*{1in}

\item What are your general impressions of the course?  Would you recommend that it continue to be offered?

\end{enumerate}


